Applying the metallicity-dependent binary fraction to double white dwarf formation: \\ Implications for LISA

### 2022-04-30 src/ms.tex
Update ms.tex
added line numbers
add linenums for journal
Update ms.tex

% Define document class
\documentclass[twocolumn]{aastex631}

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\usepackage{amsmath}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\newcommand{\cosmic}{\texttt{COSMIC}}
\newcommand{\legwork}{\texttt{LEGWORK}}

@@

@@

We discard any DWDs for which the sum of their ZAMS birth time, given by the star particle formation time, and DWD formation time is larger than the age of the star particle since the system will not have evolved long enough to become a DWD at present. We further discard any DWDs for which the lower mass WD overflows it's Roche lobe before present day, because the outcomes of these interactions are highly uncertain and their treatment is outside the scope of this work \citep[e.g., ][]{Shen2015, Kremer2017}. The separation at which the lower mass WD overflows its Roche Lobe is defined as

@@

\section{Binary evolution assumption variations}
\label{sec:model_compare}

\begin{figure}
	\includegraphics[width=0.45\textwidth]{figures/model_comp.pdf}
    \caption{The top panel shows the Galactic DWD confusion fit vs GW frequency for different binary evolution assumption variations (colors) with model F50 shown in dashed lines and model FZ shown in solid lines. The bottom panel shows the ratio of the number of DWDs orbiting in the LISA frequency band for model FZ to model F50 vs to number of DWDs orbiting in the LISA band for model F50 only. While both the height of the confusion foreground and number of LISA DWDs changes for each variation, the FZ models within each variation exhibit a constant reduction by a factor of two compared with the F50 models.}
    \label{fig:model_comp}
\end{figure}

In order to test the robustness of the reduction of the height of the Galactic DWD GW foreground when assuming a metallicity-dependent binary fraction, we repeat our analysis using three different binary evolution assumption variations. For each variation, we consider models FZ and F50 as done in the fiducial case described above. In variation $q3$, we vary the assumption for the critical mass ratios at which a RLO interaction remains stable or becomes unstable from our fiducial assumptions such that the critical mass ratio is increased to $3.0$ and thus allows stable mass transfer for more massive RLO donors. In variations $\alpha25$ and $\alpha5$, we modify the common envelope ejection efficiency to be either much less ($\alpha=0.25$) or more ($\alpha=5$) than in our fiducial assumption ($\alpha=1$). Larger common envelope ejection efficiencies lead to wider post-CE separations, while smaller ejection efficiencies either lead to closer post-CE separations or mergers where the envelope ejection fails.

@@

We extended our study to include three binary evolution assumption variations to investigate whether the DWD population reduction was robust to changes in assumptions for mass transfer stability and common envelope ejection efficiencies. While our binary evolution assumption variations change the size of the LISA-observable populations dramatically, the reduction in the size of the close Galactic DWD population and the height of the confusion foreground for models which assume a metallicity-dependent binary fraction is robust. An important consequence of a lower Galactic DWD confusion foreground is that relative to the total DWD population, more DWDs can be individually resolved because of the reduction in competing GW signals. While the number of DWDs radiating GWs in the LISA frequency band is reduced by a factor of two for model FZ relative to model F50, the number of resolved sources is less affected with a population-wide reduction of $14\%$. These results are far-reaching since the strength of the Galactic DWD confusion foreground has direct consequences on the detectability of all other LISA sources with small SNRs. An increase in resolution capability from the reduced confusion foreground can be extended to other galactic binaries that LISA will observe at these frequencies like those involving neutron stars and stellar-origin black holes, as well as other more exotic GW sources like merging supermassive black holes, extreme mass ratio inspirals, or cosmological GW backgrounds. Based on our results, we suggest that studies which employ fits to the confusion foreground based on population synthesis results consider reducing the strength of the Galactic foreground PSD by a factor of two.

>>>>

% Define document class
\documentclass[twocolumn, linenumbers]{aastex631}

\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\usepackage{amsmath}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
\newcommand{\cosmic}{\texttt{COSMIC}}
\newcommand{\legwork}{\texttt{LEGWORK}}
\usepackage{xcolor}

@@

%I did some switching around of words
\textbf{The transition between individually resolved DWDs and the confusion limited, or unresolved, DWD foreground is expected to occur near $\sim1\,\rm{mHz}$ frequencies \citep[e.g. ][]{Ruiter2010}. In the confusion limited regime, more than one binary radiates GWs in each LISA frequency bin, thus creating a superposition of signals which are unable to be disentangled. There are multiple parameters which shape the Galactic DWD foreground. Assuming the evolution of each DWD system is driven solely by GW emission, the frequency derivative scales proportionally to $f_{\rm{GW}}^{11/3}$, causing a pileup of DWDs at lower frequencies \citep{Breivik2019fg}. The spatial density of DWDs in the Galaxy  which defines the distance to each binary also impacts the foreground amplitude, since strain scales inversely with distance. Star formation history assumptions combine these two effects by assigning ages and positions to each DWD in the population which determine the present day orbital (and thus GW) frequencies as well as distances to each source.}

%\textbf{There also exists a ``transition" GW frequency at which the GW spectrum transitions from confusion limited to being able to detect individual sources \citep{Ruiter2010}. In the confusion-limited regime, greater than one binary exists per LISA frequency bin, creating a superposition of signals which are unable to be disentangled. There are multiple parameters which serve to shape the Galactic DWD foreground. Assuming the evolution of each DWD system is driven solely by GW emission, the frequency derivative scales proportionally to $f_{\rm{GW}}^{11/3}$, causing a pileup of DWDs at lower frequencies \citep{Breivik2019fg}. The distance dependence of the spatial density of DWDs in the Galaxy also impacts the foreground amplitude, since strain scales inversely with distance. Star formation history assumptions impact the foreground height, as the duration of GW orbital evolution of each system will impact contributions to strain.}

@@

We discard any DWDs for which the sum of their ZAMS birth time, given by the star particle formation time, and DWD formation time is larger than the age of the star particle since the system will not have evolved long enough to become a DWD at present. We further discard any DWDs for which the lower mass WD overflows it's Roche lobe before present day, because the outcomes of these interactions are highly uncertain and their treatment is outside the scope of this work \citep[e.g., ][]{Shen2015, Kremer2017}. \textbf{We also note that stably accreting WD binaries are unlikely to contribute appreciably to the Galactic DWD foreground due to their small mass ratios at frequencies below $1\,\rm{mHz}$ \citep{Breivik2018}.} The separation at which the lower mass WD overflows its Roche Lobe is defined as

@@

\section{Binary evolution \textbfparameter variations}
\label{sec:model_compare}

\begin{figure}
	\includegraphics[width=0.45\textwidth]{figures/model_comp.pdf}
    \caption{The top panel shows the Galactic DWD confusion fit vs GW frequency for different binary evolution \textbf{parameter variations} (colors) with model F50 shown in dashed lines and model FZ shown in solid lines. The bottom panel shows the ratio of the number of DWDs orbiting in the LISA frequency band for model FZ to model F50 vs to number of DWDs orbiting in the LISA band for model F50 only. While both the height of the confusion foreground and number of LISA DWDs changes for each variation, the FZ models within each variation exhibit a constant reduction by a factor of two compared with the F50 models.}
    \label{fig:model_comp}
\end{figure}

In order to test the robustness of the reduction of the height of the Galactic DWD GW foreground when assuming a metallicity-dependent binary fraction, we repeat our analysis using three different binary evolution \textbf{parameter variations}. For each variation, we consider models FZ and F50 as done in the fiducial case described above. In variation $q3$, we vary the assumption for the critical mass ratios at which a RLO interaction remains stable or becomes unstable from our fiducial assumptions such that the critical mass ratio is increased to $3.0$ and thus allows stable mass transfer for more massive RLO donors. In variations $\alpha25$ and $\alpha5$, we modify the common envelope ejection efficiency to be either much less ($\alpha=0.25$) or more ($\alpha=5$) than in our fiducial assumption ($\alpha=1$). Larger common envelope ejection efficiencies lead to wider post-CE separations, while smaller ejection efficiencies either lead to closer post-CE separations or mergers where the envelope ejection fails.

@@

We extended our study to include three binary evolution \textbf{parameter variations} to investigate whether the DWD population reduction was robust to changes in assumptions for mass transfer stability and common envelope ejection efficiencies. While our binary evolution \textbf{parameter variations} change the size of the LISA-observable populations dramatically, the reduction in the size of the close Galactic DWD population and the height of the confusion foreground for models which assume a metallicity-dependent binary fraction is robust. An important consequence of a lower Galactic DWD confusion foreground is that relative to the total DWD population, more DWDs can be individually resolved because of the reduction in competing GW signals. While the number of DWDs radiating GWs in the LISA frequency band is reduced by a factor of two for model FZ relative to model F50, the number of resolved sources is less affected with a population-wide reduction of $14\%$. These results are far-reaching since the strength of the Galactic DWD confusion foreground has direct consequences on the detectability of all other LISA sources with small SNRs. An increase in resolution capability from the reduced confusion foreground can be extended to other galactic binaries that LISA will observe at these frequencies like those involving neutron stars and stellar-origin black holes, as well as other more exotic GW sources like merging supermassive black holes, extreme mass ratio inspirals, or cosmological GW backgrounds. Based on our results, we suggest that studies which employ fits to the confusion foreground based on population synthesis results consider reducing the strength of the Galactic foreground PSD by a factor of two.
